---
title: "CMC for Presentation"
author: "Bo Wang"
date: "December 20, 2016"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, include=FALSE}
library(dplyr)
library(plyr)
library(ggplot2)
library("fmsb")
```

1. Data: CMC-Interposition arthroplasty, intercarpal or carpometacarpal joints (25447) and Arthroplasty with prosthetic replacement, trapezium (25445). 

```{r, echo=FALSE, include=FALSE}
load(file="CMC.RDA")
load(file = "CMC05_06.RDA")
load(file = "CMC07.RDA")
load(file = "CMC08.RDA")
load(file = "CMC09.RDA")
load(file = "CMC10.RDA")
load(file = "CMC11.RDA")
load(file = "CMC12.RDA")
load(file = "CMC13.RDA")
load(file = "CMC14.RDA")
load(file = "CMC15.RDA")
```

2. Basic summary of CMC data

Number of patients and variables overall
```{r, echo=FALSE, include=FALSE}
dim<-dim(CMC)
names(dim)<-c("patients", "variables")
dim
```

Number of patients and variables for each year
```{r, echo=FALSE, include=FALSE}
dim05_06<-dim(CMC05_06)
dim07<-dim(CMC07)
dim08<-dim(CMC08)
dim09<-dim(CMC09)
dim10<-dim(CMC10)
dim11<-dim(CMC11)
dim12<-dim(CMC12)
dim13<-dim(CMC13)
dim14<-dim(CMC14)
dim15<-dim(CMC15)
dims<-c(dim05_06, dim07, dim08, dim09, dim10, dim11, dim12, dim13, dim14, dim15)
dims<-as.data.frame(matrix(dims, nrow = 2))
rownames(dims)<-c("patients", "variables")
colnames(dims)<-c("2005-06", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")
dims
```



3. Complications

Number of unique patients with complications (1.6%)
```{r, echo=FALSE}
### function to assign Y/N complication column 
complicationOrNot<-function(df){
  Complication<-vector()
  complications<-c("SUPINFEC", "WNDINFD", "ORGSPCSSI",
        "DEHIS", "OUPNEUMO", "REINTUB", "PULEMBOL","FAILWEAN", 
        "RENAINSF", "OPRENAFL", "URNINFEC", "CNSCVA",
        "CDARREST", "CDMI", "OTHBLEED", "OTHDVT", "OTHSYSEP", 
        "OTHSESHOCK", "OTHGRAFL", "CNSCOMA", "NEURODEF",
        "OTHCDIFF",   ## "OTHCDIFF" added in 2015, no cases
        "RETORRELATED", "READMRELATED1", "DOPERTOD")
for(i in 1:nrow(df)) {if((all(df[,complications][i,]=="No" 
                          | df[,complications][i,]=="No Complication"
                          | df[,complications][i,]=="NULL"
                          | df[,complications][i,]== -99
                          | df[,complications][i,]=="Unknown"
                          | is.na(df[,complications][i,])))==TRUE){
                 Complication[i]<-"FALSE"
                 } else {
                 Complication[i]<-"TRUE"}}
  df<-cbind(df, Complication)
}
#CMCaddComplication<-complicationOrNot(CMC)
#CMCaddComplication[CMC$DOPERTOD==29,]$Complication<-TRUE

load("CMCaddComplication.RDA")

table(CMCaddComplication$Complication)
```

Distribution of complications
```{r, echo=FALSE}
CMCwithComplication<-CMCaddComplication[CMCaddComplication$Complication==TRUE,]
CMCwoComplication<-CMCaddComplication[CMCaddComplication$Complication==FALSE,]
complicationTables<-list(
table(CMCwithComplication$SUPINFEC),
table(CMCwithComplication$WNDINFD),
table(CMCwithComplication$ORGSPCSSI),
table(CMCwithComplication$DEHIS),
table(CMCwithComplication$OUPNEUMO),
table(CMCwithComplication$PULEMBOL),
table(CMCwithComplication$REINTUB),
table(CMCwithComplication$FAILWEAN),
table(CMCwithComplication$CDARREST),
table(CMCwithComplication$CDMI),
table(CMCwithComplication$RENAINSF),
table(CMCwithComplication$OPRENAFL),
table(CMCwithComplication$CNSCVA),
table(CMCwithComplication$OTHGRAFL),
table(CMCwithComplication$CNSCOMA),
table(CMCwithComplication$NEURODEF),
table(CMCwithComplication$OTHDVT),
table(CMCwithComplication$OTHSYSEP),
table(CMCwithComplication$OTHSESHOCK),
table(CMCwithComplication$RETORRELATED),
table(CMCwithComplication$OTHBLEED),
table(CMCwithComplication$URNINFEC),
table(CMCwithComplication$READMRELATED1)
)

compl1<-c("SUPINFEC", "WNDINFD", 
        "ORGSPCSSI",
        "DEHIS", "OUPNEUMO","PULEMBOL", 
        "REINTUB", "FAILWEAN", 
        "CDARREST","CDMI", "RENAINSF", "OPRENAFL",
        "CNSCVA", "OTHGRAFL",
        "CNSCOMA","NEURODEF", "OTHDVT",
        "OTHSYSEP", "OTHSESHOCK",
        "RETORRELATED", "OTHBLEED",
        "URNINFEC", "READMRELATED1")
complicationdfs<-lapply(complicationTables, tbToDf)
names(complicationdfs)<-compl1
complicationdfs
```


RETURNOR/READMISSION
```{r, echo=FALSE, include=FALSE}
load(file = "READMISSION.RDA")
load(file = "RETURNOR.RDA")
READMISSION
RETURNOR
```



4. 2*2 Tables and Chi-squared analysis (categorical variable v. Complication (Y/N))

```{r}
##function to get rid of all 0 rows in a table
ridZero<-function(tb){
  num<-1
  while(num<=dim(tb)[1]){
    if(all(tb[num,]==0)==TRUE){
      tb<-tb[-num,]
      num<-num
    }else{
      tb<-tb
      num<-num+1
    }
  }
  tb
}

## Fisher's exact or chi-squared 
## https://www.graphpad.com/guides/prism/6/statistics/index.htm?stat_chi-square_or_fishers_test.htm
```

SEX check
```{r, echo=FALSE}
SEXtb<-table(CMCaddComplication[, c("SEX", "Complication")])
SEXtb
SEXtb<-SEXtb[-3,]
SEXtb
SEXchisq<-chisq.test(SEXtb, simulate.p.value = TRUE)
SEXchisq$p.value
```

RACE check
```{r, echo=FALSE}
RACEtb<-table(CMCaddComplication[, c("RACE", "Complication")])
RACEtb
RACEtb[5,]<-c(495, 8)
RACEtb<-RACEtb[c(-1,-4),]
RACEtb
RACEchisq<-chisq.test(RACEtb, simulate.p.value = TRUE)
RACEchisq$p.value
```

INOUT check
```{r, echo=FALSE}
INOUTtb<-table(CMCaddComplication[, c("INOUT", "Complication")])
INOUTtb<-INOUTtb[-3,]
INOUTtb
INOUTchisq<-chisq.test(INOUTtb, simulate.p.value = TRUE)
INOUTchisq$p.value
INOUTtb[,1]/(INOUTtb[,1]+INOUTtb[,2])
```


TRANSFUS
```{r, echo=FALSE}
TRANSFUStb<-table(CMCaddComplication[, c("TRANSFUS", "Complication")])
TRANSFUStb<-ridZero(TRANSFUStb)
TRANSFUStb
TRANSFUSchisq<-chisq.test(TRANSFUStb, simulate.p.value = TRUE)
TRANSFUSchisq$p.value
```

PROPER30
```{r, echo=FALSE}
PROPER30tb<-table(CMCaddComplication[, c("PROPER30", "Complication")])
PROPER30tb<-ridZero(PROPER30tb)
PROPER30tb<-PROPER30tb[-2,]
PROPER30chisq<-chisq.test(PROPER30tb, simulate.p.value = TRUE)
PROPER30chisq$p.value
```

TRANST
```{r, echo=FALSE}
TRANSTtb<-table(CMCaddComplication[, c("TRANST", "Complication")])
TRANSTtb<-ridZero(TRANSTtb)
TRANSTtb
TRANSTchisq<-chisq.test(TRANSTtb, simulate.p.value = TRUE)
TRANSTchisq$p.value

TRANSTtb2<-TRANSTtb[c(1,4),]
TRANSTtb2
TRANSTchisq2<-chisq.test(TRANSTtb2, simulate.p.value = TRUE)
TRANSTchisq2$p.value
```

ATTEND check
```{r, echo=FALSE}
ATTENDtb<-table(CMCaddComplication[, c("ATTEND", "Complication")])
ATTENDtb<-ridZero(ATTENDtb)
ATTENDtb
ATTENDtb2<-ATTENDtb[c(-3,-5),]
ATTENDtb2
ATTENDchisq2<-chisq.test(ATTENDtb2, simulate.p.value = TRUE)
ATTENDchisq2$p.value
```


Resident vs No resident check
```{r}
ATTEND<-matrix(c((498+84),9,238,6),nrow = 2, byrow = TRUE)
rownames(ATTEND)<-c("No resident", "Resident")
colnames(ATTEND)<-c("FALSE", "TRUE")
ATTEND<-as.table(ATTEND)
ATTEND
ATTENDchisq<-chisq.test(ATTEND, simulate.p.value = TRUE)
ATTENDchisq$p.value
ATTEND[,1]/(ATTEND[,1]+ATTEND[,2])
```


70+ or not
```{r, echo=FALSE}
CMC70plus<-CMCaddComplication[CMCaddComplication$AGE>=70,]
CMCbelow70<-CMCaddComplication[CMCaddComplication$AGE<70,]
table(CMCbelow70$Complication)
table(CMC70plus$Complication)
seventyPLUS<-matrix(c(688,16, 2611, 29), nrow = 2, byrow = TRUE)
rownames(seventyPLUS)<-c("Yes", "No")
colnames(seventyPLUS)<-c("FALSE", "TRUE")
seventyPLUS<-as.table(seventyPLUS)
seventyPLUS
seventyPLUSchisq<-chisq.test(seventyPLUS, simulate.p.value = TRUE)
seventyPLUSchisq$p.value
seventyPLUS[,1]/(seventyPLUS[,1]+seventyPLUS[,2])
```


65+ or not
```{r, echo=FALSE}
CMC65plus<-CMCaddComplication[CMCaddComplication$AGE>=65,]
CMCbelow65<-CMCaddComplication[CMCaddComplication$AGE<65,]
table(CMCbelow65$Complication)
table(CMC65plus$Complication)
sixtyfivePLUS<-matrix(c(1321,25, 1978, 20), nrow = 2, byrow = TRUE)
rownames(sixtyfivePLUS)<-c("Yes", "No")
colnames(sixtyfivePLUS)<-c("FALSE", "TRUE")
sixtyfivePLUS<-as.table(sixtyfivePLUS)
sixtyfivePLUS
sixtyfivePLUSchisq<-chisq.test(sixtyfivePLUS, simulate.p.value = TRUE)
sixtyfivePLUSchisq$p.value
sixtyfivePLUS[,1]/(sixtyfivePLUS[,1]+sixtyfivePLUS[,2])
```


60+ or not
```{r, echo=FALSE}
CMC60plus<-CMCaddComplication[CMCaddComplication$AGE>=60,]
CMCbelow60<-CMCaddComplication[CMCaddComplication$AGE<60,]
table(CMCbelow60$Complication)
table(CMC60plus$Complication)
sixtyPLUS<-matrix(c(1994,29, 1305, 16), nrow = 2, byrow = TRUE)
rownames(sixtyPLUS)<-c("Yes", "No")
colnames(sixtyPLUS)<-c("FALSE", "TRUE")
sixtyPLUS<-as.table(sixtyPLUS)
sixtyPLUS
sixtyPLUSchisq<-chisq.test(sixtyPLUS, simulate.p.value = TRUE)
sixtyPLUSchisq$p.value
sixtyPLUS[,1]/(sixtyPLUS[,1]+sixtyPLUS[,2])
```


ANESTHES check
```{r, echo=FALSE}
ANESTHEStb<-table(CMCaddComplication[, c("ANESTHES", "Complication")])
ANESTHEStb<-ridZero(ANESTHEStb)
ANESTHEStb
ANESTHESchisq<-chisq.test(ANESTHEStb, simulate.p.value = TRUE)
ANESTHESchisq$p.value
ANESTHEStb[,2]/(ANESTHEStb[,1]+ANESTHEStb[,2])
ANESTHEStb2<-ANESTHEStb[c(2,7,11),]
ANESTHEStb2
ANESTHESchisq2<-chisq.test(ANESTHEStb2, simulate.p.value = TRUE)
ANESTHESchisq2$p.value
ANESTHEStb2[,2]/(ANESTHEStb2[,1]+ANESTHEStb2[,2])
```

General anesthesia vs all others check
```{r}
ANE2<-matrix(c(2268, 32, (1+6+22+4+7+553+2+429), 13), nrow = 2, byrow = TRUE)
rownames(ANE2)<-c("General", "Others")
colnames(ANE2)<-c("FALSE", "TRUE")
ANE2<-as.table(ANE2)
ANE2
ANE2chisq<-chisq.test(ANE2, simulate.p.value = TRUE)
ANE2chisq$p.value
ANE2[,1]/(ANE2[,1]+ANE2[,2])
```

SURGSPEC
```{r, echo=FALSE}
SURGSPECtb<-table(CMCaddComplication[, c("SURGSPEC", "Complication")])
SURGSPECtb<-ridZero(SURGSPECtb)
SURGSPECtb
SURGSPECchisq<-chisq.test(SURGSPECtb, simulate.p.value = TRUE)
SURGSPECchisq$p.value
SURGSPECtb2<-SURGSPECtb[c(3,5),]
SURGSPECtb2
SURGSPECchisq2<-chisq.test(SURGSPECtb2, simulate.p.value = TRUE)
SURGSPECchisq2$p.value
SURGSPECtb2[,1]/(SURGSPECtb2[,1]+SURGSPECtb2[,2])
```

DIABETES check
```{r, echo=FALSE}
DIABETEStb<-table(CMCaddComplication[, c("DIABETES", "Complication")])
DIABETEStb<-ridZero(DIABETEStb)
DIABETEStb
DIABETESchisq<-chisq.test(DIABETEStb, simulate.p.value = TRUE)
DIABETESchisq$p.value
DIABETEStb[,2]/(DIABETEStb[,1]+DIABETEStb[,2])
CMCaddComplication$DIABETES<-factor(CMCaddComplication$DIABETES, levels = c("NO", "INSULIN", "ORAL","NON-INSULIN"))
```

No Diabetes v Yes Diabetes 
```{r, echo=FALSE}
DIA<-matrix(c(2961, 43, 338, 8), nrow = 2, byrow = TRUE)
rownames(DIA)<-c("No", "Yes")
colnames(DIA)<-c("FALSE", "TRUE")
DIA<-as.table(DIA)
DIA
DIAchisq<-chisq.test(DIA, simulate.p.value = TRUE)
DIAchisq$p.value

```

No Diabetes v Insulin check
```{r, echo=FALSE}
DIA2<-matrix(c(2961, 37, 110, 5), nrow = 2, byrow = TRUE)
rownames(DIA2)<-c("No", "Insulin")
colnames(DIA2)<-c("FALSE", "TRUE")
DIA2<-as.table(DIA2)
DIA2
DIA2chisq<-chisq.test(DIA2, simulate.p.value = TRUE)
DIA2chisq$p.value
```

No Insulin v Insulin check
```{r, echo=FALSE}
DIA3<-matrix(c(3189, 40, 110, 5), nrow = 2, byrow = TRUE)
rownames(DIA3)<-c("No insulin", "Insulin")
colnames(DIA3)<-c("FALSE", "TRUE")
DIA3<-as.table(DIA3)
DIA3
DIA3chisq<-chisq.test(DIA3, simulate.p.value = TRUE)
DIA3chisq$p.value
DIA3[,2]/(DIA3[,1]+DIA3[,2])
```

WOUND_CLOSURE check
```{r, echo=FALSE}
WOUND_CLOSUREtb<-table(CMCaddComplication[, c("WOUND_CLOSURE", "Complication")])
WOUND_CLOSUREtb<-ridZero(WOUND_CLOSUREtb)
WOUND_CLOSUREtb
WOUND_CLOSUREchisq<-chisq.test(WOUND_CLOSUREtb, simulate.p.value = TRUE)
WOUND_CLOSUREchisq$p.value
WOUND_CLOSUREtb[,1]/(WOUND_CLOSUREtb[,1]+WOUND_CLOSUREtb[,2])

WOUND_CLOSURE2<-matrix(c(1537, 26, 2, 0), nrow = 2, byrow = TRUE)
rownames(WOUND_CLOSURE2)<-c("Closed", "Not closed")
colnames(WOUND_CLOSURE2)<-c("FALSE", "TRUE")
WOUND_CLOSURE2<-as.table(WOUND_CLOSURE2)
WOUND_CLOSURE2
WOUND_CLOSURE2chisq<-chisq.test(WOUND_CLOSURE2, simulate.p.value = TRUE)
WOUND_CLOSURE2chisq$p.value
WOUND_CLOSURE2[,1]/(WOUND_CLOSURE2[,1]+WOUND_CLOSURE2[,2])
```


ETHNICITY_HISPANIC check
```{r, echo=FALSE}
ETHNICITY_HISPANICtb<-table(CMCaddComplication[, c("ETHNICITY_HISPANIC", "Complication")])
ETHNICITY_HISPANICtb<-ridZero(ETHNICITY_HISPANICtb)
ETHNICITY_HISPANICtb<-ETHNICITY_HISPANICtb[1:2,]
ETHNICITY_HISPANICtb
ETHNICITY_HISPANICchisq<-chisq.test(ETHNICITY_HISPANICtb, simulate.p.value = TRUE)
ETHNICITY_HISPANICchisq$p.value
ETHNICITY_HISPANICtb[,1]/(ETHNICITY_HISPANICtb[,1]+ETHNICITY_HISPANICtb[,2])
```


DIALYSIS check
```{r, echo=FALSE}
DIALYSIStb<-table(CMCaddComplication[, c("DIALYSIS", "Complication")])
DIALYSIStb<-ridZero(DIALYSIStb)
DIALYSIStb
DIALYSISchisq<-chisq.test(DIALYSIStb, simulate.p.value = TRUE)
DIALYSISchisq$p.value
DIALYSIStb[,1]/(DIALYSIStb[,1]+DIALYSIStb[,2])
```

SMOKE check
```{r, echo=FALSE}
SMOKEtb<-table(CMCaddComplication[, c("SMOKE", "Complication")])
SMOKEtb<-ridZero(SMOKEtb)
SMOKEtb
SMOKEchisq<-chisq.test(SMOKEtb, simulate.p.value = TRUE)
SMOKEchisq$p.value
```

ETOH(alcohol use) check
```{r, echo=FALSE}
ETOHtb<-table(CMCaddComplication[, c("ETOH", "Complication")])
ETOHtb<-ridZero(ETOHtb)
ETOHtb<-ETOHtb[-3,]
ETOHtb
ETOHchisq<-chisq.test(ETOHtb, simulate.p.value = TRUE)
ETOHchisq$p.value
```

CVA(cerebrovascular accident) check
```{r, echo=FALSE}
CVAtb<-table(CMCaddComplication[, c("CVA", "Complication")])
CVAtb<-ridZero(CVAtb)
CVAtb<-CVAtb[-3,]
CVAtb
CVAchisq<-chisq.test(CVAtb, simulate.p.value = TRUE)
CVAchisq$p.value

CVANOtb<-table(CMCaddComplication[, c("CVANO", "Complication")])
CVANOtb<-ridZero(CVANOtb)
CVANOtb<-CVANOtb[-3,]
CVANOtb
CVANOchisq<-chisq.test(CVANOtb, simulate.p.value = TRUE)
CVANOchisq$p.value
```

HXTIA (transient ischemic attack) check
```{r, echo=FALSE}
HXTIAtb<-table(CMCaddComplication[, c("HXTIA", "Complication")])
HXTIAtb<-ridZero(HXTIAtb)
HXTIAtb<-HXTIAtb[-3,]
HXTIAtb
HXTIAchisq<-chisq.test(HXTIAtb, simulate.p.value = TRUE)
HXTIAchisq$p.value
```

CVA or TIA check
```{r, echo=FALSE}
CVATIAOrNot<-function(df){
  CVATIA<-vector()
  temp<-df[, c("CVA", "CVANO", "HXTIA")]
for(i in 1:nrow(df)) {if(any(temp[i,]=="Yes")){
                 CVATIA[i]<-"Yes"
                 } else {
                 CVATIA[i]<-"No"}}
  df<-cbind(df, CVATIA)
}

CVAY<-CMC[CMC$CVA=="Yes",]$CASEID
CVAY<-CVAY[1:10]
CVAN<-CMC[CMC$CVA=="No",]$CASEID
CVAN<-CVAN[1:826]
CVAID<-c(CVAY, CVAN)

CVANOY<-CMC[CMC$CVANO=="Yes",]$CASEID
CVANOY<-CVANOY[1:13]
CVANON<-CMC[CMC$CVANO=="No",]$CASEID
CVANON<-CVANON[1:824]
CVANOID<-c(CVANOY, CVANON)

TIAY<-CMC[CMC$HXTIA=="Yes",]$CASEID
TIAY<-TIAY[1:34]
TIAN<-CMC[CMC$HXTIA=="No",]$CASEID
TIAN<-TIAN[1:802]
TIAID<-c(TIAY, TIAN)

Y<-Reduce(union, list(CVAY, CVANOY, TIAY))
N<-Reduce(union, list(CVAN, CVANON, TIAN))

CMCwithCVATIAY<-CMCaddComplication[CMCaddComplication$CASEID %in% Y,]
CMCwithCVATIAN<-CMCaddComplication[CMCaddComplication$CASEID %in% N,]

CMCwithCVATIAYtb<-table(CMCwithCVATIAY$Complication)
CMCwithCVATIANtb<-table(CMCwithCVATIAN$Complication)

CVATIA<-matrix(c(821,15,48,1),nrow = 2, byrow = TRUE)
rownames(CVATIA)<-c("No", "Yes")
colnames(CVATIA)<-c("FALSE", "TRUE")
CVATIA<-as.table(CVATIA)
CVATIA
CVATIAchisq<-chisq.test(CVATIA, simulate.p.value = TRUE)
CVATIAchisq$p.value
```


CPT check
```{r, echo=FALSE}
CPTtb<-table(CMCaddComplication[, c("CPT", "Complication")])
CPTtb<-ridZero(CPTtb)
CPTtb
CPTchisq<-chisq.test(CPTtb, simulate.p.value = TRUE)
CPTchisq$p.value
```

HYPERMED check
```{r, echo=FALSE}
HYPERMEDtb<-table(CMCaddComplication[, c("HYPERMED", "Complication")])
HYPERMEDtb<-ridZero(HYPERMEDtb)
HYPERMEDtb
HYPERMEDchisq<-chisq.test(HYPERMEDtb, simulate.p.value = TRUE)
HYPERMEDchisq$p.value
```

WNDCLAS check
```{r, echo=FALSE}
WNDCLAStb<-table(CMCaddComplication[, c("WNDCLAS", "Complication")])
WNDCLAStb<-ridZero(WNDCLAStb)
WNDCLAStb
WNDCLASchisq<-chisq.test(WNDCLAStb, simulate.p.value = TRUE)
WNDCLASchisq$p.value
```

ASACLAS check
```{r, echo=FALSE}
ASACLAStb<-table(CMCaddComplication[, c("ASACLAS", "Complication")])
ASACLAStb<-ridZero(ASACLAStb)
ASACLAStb<-ASACLAStb[-5,]
ASACLAStb
ASACLASchisq<-chisq.test(ASACLAStb, simulate.p.value = TRUE)
ASACLASchisq$p.value
ASACLAStb[,2]/(ASACLAStb[,1]+ASACLAStb[,2])

ASAtemp<-matrix(c(2284,25, 959,20),nrow = 2, byrow = TRUE)
rownames(ASAtemp)<-c("onetwo", "threefour")
colnames(ASAtemp)<-c("FALSE", "TRUE")
ASAtemp<-as.table(ASAtemp)
ASAtemp
ASAtempchisq<-chisq.test(ASAtemp, simulate.p.value = TRUE)
ASAtempchisq$p.value
```

BLEEDDIS check
```{r, echo=FALSE}
BLEEDDIStb<-table(CMCaddComplication[, c("BLEEDDIS", "Complication")])
BLEEDDIStb<-ridZero(BLEEDDIStb)
BLEEDDIStb
BLEEDDISchisq<-chisq.test(BLEEDDIStb, simulate.p.value = TRUE)
BLEEDDISchisq$p.value
BLEEDDIStb[,2]/(BLEEDDIStb[,1]+BLEEDDIStb[,2])
```

HXCHF (congestive heart failure) check
```{r, echo=FALSE}
HXCHFtb<-table(CMCaddComplication[, c("HXCHF", "Complication")])
HXCHFtb<-ridZero(HXCHFtb)
HXCHFtb
HXCHFchisq<-chisq.test(HXCHFtb, simulate.p.value = TRUE)
HXCHFchisq$p.value
```

HXPVD (peripheral vascular disease) check
```{r, echo=FALSE}
HXPVDtb<-table(CMCaddComplication[, c("HXPVD", "Complication")])
HXPVDtb<-ridZero(HXPVDtb)
HXPVDtb<-HXPVDtb[-3,]
HXPVDtb
HXPVDchisq<-chisq.test(HXPVDtb, simulate.p.value = TRUE)
HXPVDchisq$p.value
```

FNSTATUS2 (Functional status before surgery) check
```{r, echo=FALSE}
FNSTATUS2tb<-table(CMCaddComplication[, c("FNSTATUS2", "Complication")])
FNSTATUS2tb
FNSTATUS2tb<-ridZero(FNSTATUS2tb)
FNSTATUS2tb<-FNSTATUS2tb[-3,]
FNSTATUS2tb
FNSTATUS2chisq<-chisq.test(FNSTATUS2tb, simulate.p.value = TRUE)
FNSTATUS2chisq$p.value
```

STEROID check
```{r, echo=FALSE}
STEROIDtb<-table(CMCaddComplication[, c("STEROID", "Complication")])
STEROIDtb<-ridZero(STEROIDtb)
STEROIDtb
STEROIDchisq<-chisq.test(STEROIDtb, simulate.p.value = TRUE)
STEROIDchisq$p.value
STEROIDtb[,1]/(STEROIDtb[,1]+STEROIDtb[,2])
```


PRVPCS (previous cardiac surgery) check
```{r, echo=FALSE}
PRVPCStb<-table(CMCaddComplication[, c("PRVPCS", "Complication")])
PRVPCStb<-ridZero(PRVPCStb)
PRVPCStb<-PRVPCStb[-3,]
PRVPCStb
PRVPCSchisq<-chisq.test(PRVPCStb, simulate.p.value = TRUE)
PRVPCSchisq$p.value
```


HXANGINA check
```{r, echo=FALSE}
HXANGINAtb<-table(CMCaddComplication[, c("HXANGINA", "Complication")])
HXANGINAtb<-ridZero(HXANGINAtb)
HXANGINAtb<-HXANGINAtb[-3,]
HXANGINAtb
HXANGINAchisq<-chisq.test(HXANGINAtb, simulate.p.value = TRUE)
HXANGINAchisq$p.value
```


CHEMO
```{r, echo=FALSE}
CHEMOtb<-table(CMCaddComplication[, c("CHEMO", "Complication")])
CHEMOtb<-ridZero(CHEMOtb)
CHEMOtb<-CHEMOtb[-3,]
CHEMOtb
CHEMOchisq<-chisq.test(CHEMOtb, simulate.p.value = TRUE)
CHEMOchisq$p.value
```

RADIO
```{r, echo=FALSE}
RADIOtb<-table(CMCaddComplication[, c("RADIO", "Complication")])
RADIOtb<-ridZero(RADIOtb)
RADIOtb<-RADIOtb[-3,]
RADIOtb
RADIOchisq<-chisq.test(RADIOtb, simulate.p.value = TRUE)
RADIOchisq$p.value
```


CHEMO or RADIO check
```{r, echo=FALSE}
CHEMOY<-CMC[CMC$CHEMO=="Yes",]$CASEID
CHEMOY<-CHEMOY[1]
CHEMON<-CMC[CMC$CHEMO=="No",]$CASEID
CHEMON<-CHEMON[1:835]
CHEMOID<-c(CHEMOY, CHEMON)

RADIOY<-CMC[CMC$RADIO=="Yes",]$CASEID
RADION<-CMC[CMC$RADIO=="No",]$CASEID
RADION<-RADION[1:836]
Y<-Reduce(union, list(CHEMOY))
N<-Reduce(union, list(CHEMON, RADION))

CMCwithCHEMORADIOY<-CMCaddComplication[CMCaddComplication$CASEID %in% Y,]
CMCwithCHEMORADION<-CMCaddComplication[CMCaddComplication$CASEID %in% N,]

table(CMCwithCHEMORADIOY$Complication)
table(CMCwithCHEMORADION$Complication)

CHEMORADIO<-matrix(c(821,15,1,0),nrow = 2, byrow = TRUE)
rownames(CHEMORADIO)<-c("No", "Yes")
colnames(CHEMORADIO)<-c("FALSE", "TRUE")
CHEMORADIO<-as.table(CHEMORADIO)
CHEMORADIO
CHEMORADIOchisq<-chisq.test(CHEMORADIO, simulate.p.value = TRUE)
CHEMORADIOchisq$p.value
```


PRVPCI (all but year 2015) (previous Percutaneous Coronary Intervention) check
```{r, echo=FALSE}
PRVPCItb<-table(CMCaddComplication[, c("PRVPCI", "Complication")])
PRVPCItb<-ridZero(PRVPCItb)
PRVPCItb<-PRVPCItb[-3,]
PRVPCItb
PRVPCIchisq<-chisq.test(PRVPCItb, simulate.p.value = TRUE)
PRVPCIchisq$p.value
PRVPCItb[,1]/(PRVPCItb[,1]+PRVPCItb[,2])

```

TRANSFUS
```{r, echo=FALSE}
TRANSFUStb<-table(CMCaddComplication[, c("TRANSFUS", "Complication")])
TRANSFUStb
```



Predictors I didn't find in dataset (PUF):

Intra-articular fracture

Associated ORIF ulna styloid

Current alcohol use

Preoperative function - FNSTATUS1/FNSTATUS2
All year but 2015 has both variables, 2015 only has FNSTATUS2
```{r}
table(CMCaddComplication$FNSTATUS1)
```

Chronic pulmonary obstructive disease

Peripheral vascular disease

Cardiac disease

Cerebrovascular accident or transient ischemic attack



5. T-tests (Continuous variable v. Complication (Y/N)))

WORKRVU
```{r, echo=FALSE}
WORKRVUtt<-t.test(WORKRVU~Complication, data = CMCaddComplication)
WORKRVUtt$p.value
```

AGE check
```{r}
AGEtt<-t.test(AGE~Complication, data = CMCaddComplication)
AGEtt
sd(CMCwoComplication$AGE)
range(CMCwoComplication$AGE)
sd(CMCwithComplication$AGE)
range(CMCwithComplication$AGE)
```

```{r, include=FALSE}
over65<-function(df){
  sixtyfive<-vector()
for(i in 1:nrow(df)) {
  if(df$AGE[i]>65){
  sixtyfive[i]<-TRUE
  } else{
  sixtyfive[i]<-FALSE
  }
  }
  df<-cbind(df, sixtyfive)
  df
}

CMCaddComplication<-over65(CMCaddComplication)
table(CMCaddComplication[,c("sixtyfive", "Complication")])
chisq.test(table(CMCaddComplication[,c("sixtyfive", "Complication")]))

ageGroup<-function(df){
  ageGp<-vector()
for(i in 1:nrow(df)) {
  if(df$AGE[i]<=50){
  ageGp[i]<-"<50"
  } else if(df$AGE[i]<=60){
  ageGp[i]<-"51-60"
  } else if(df$AGE[i]<=70){
  ageGp[i]<-"61-70"
  } else if(df$AGE[i]<=80){
  ageGp[i]<-"71-80"
  } else{
  ageGp[i]<-"81-90"
  }
  }
  df<-cbind(df, ageGp)
  df
}

chisq.test(table(CMCaddComplication[,c("ageGp", "Complication")]))
```

BMI check
```{r, echo=FALSE}
CMCwHeight<-filter(CMCaddComplication, HEIGHT>=0)
CMCwHeightWeight<-filter(CMCwHeight, WEIGHT>=0)
dim(CMCwHeightWeight)
CMCwHeightWeight$BMI<-CMCwHeightWeight$WEIGHT*0.45/(CMCwHeightWeight$HEIGHT*0.025)^2

BMItt<-t.test(BMI~Complication, data = CMCwHeightWeight)
BMItt$p.value

sd(CMCwHeightWeight[CMCwHeightWeight$Complication=="TRUE",]$BMI)
sd(CMCwHeightWeight[CMCwHeightWeight$Complication=="FALSE",]$BMI)


bmiGroup<-function(df){
  bmiGp<-vector()
for(i in 1:nrow(df)) {
  if(df$BMI[i]<=18.5){
  bmiGp[i]<-"underweight"
  } else if(df$BMI[i]<=25){
  bmiGp[i]<-"normal"
  } else if(df$BMI[i]<=30){
  bmiGp[i]<-"overweight"
  } else{
  bmiGp[i]<-"obese"
  }
  }
  df<-cbind(df, bmiGp)
  df
}

CMCaddComplication$BMI<-CMCaddComplication$WEIGHT*0.45/(CMCaddComplication$HEIGHT*0.025)^2
CMCaddComplication<-bmiGroup(CMCaddComplication)
chisq.test(table(CMCaddComplication[,c("bmiGp", "Complication")]))
```

PRPT (check)
```{r}
CMCwPRPT<-CMCaddComplication[CMCaddComplication$PRPT!=-99,]
dim(CMCwPRPT)
table(CMCwPRPT$Complication)
```

PRINR check
```{r}
CMCwPRINR<-CMCaddComplication[CMCaddComplication$PRINR!=-99,]
dim(CMCwPRINR)
table(CMCwPRINR$Complication)
PRINRtt<-t.test(PRINR~Complication, data = CMCwPRINR)
PRINRtt$p.value

sd(CMCwPRINR[CMCwPRINR$Complication=="TRUE",]$PRINR)
sd(CMCwPRINR[CMCwPRINR$Complication=="FALSE",]$PRINR)
```

PRWBC check
```{r}
CMCwPRWBC<-filter(CMCaddComplication, PRWBC>=0)
dim(CMCwPRWBC)
PRWBCtt<-t.test(PRWBC~Complication, data = CMCwPRWBC)
PRWBCtt$p.value

sd(CMCwPRWBC[CMCwPRWBC$Complication=="TRUE",]$PRWBC)
sd(CMCwPRWBC[CMCwPRWBC$Complication=="FALSE",]$PRWBC)
```

PRCREAT check
```{r}
CMCwPRCREAT<-filter(CMCaddComplication, PRCREAT>=0)
dim(CMCwPRCREAT)
PRCREATtt<-t.test(PRCREAT~Complication, data = CMCwPRCREAT)
PRCREATtt$p.value

sd(CMCwPRCREAT[CMCwPRCREAT$Complication=="TRUE",]$PRCREAT)
sd(CMCwPRCREAT[CMCwPRCREAT$Complication=="FALSE",]$PRCREAT)
```

PRHCT (Pre-operative hematocrit) check
```{r}
CMCwPRHCT<-filter(CMCaddComplication, PRHCT>=0)
dim(CMCwPRHCT)
PRHCTtt<-t.test(PRHCT~Complication, data = CMCwPRHCT)
PRHCTtt
sd(CMCwPRHCT[CMCwPRHCT$Complication=="TRUE",]$PRHCT)
sd(CMCwPRHCT[CMCwPRHCT$Complication=="FALSE",]$PRHCT)
```

OPTIME check
```{r}
CMCwOPTIME<-filter(CMCaddComplication, OPTIME>=0)
CMCwOPTIME$OPTIME<-as.numeric(CMCwOPTIME$OPTIME)
dim(CMCwOPTIME)
OPTIMEtt<-t.test(OPTIME~Complication, data = CMCwOPTIME)
OPTIMEtt$p.value
sd(CMCwOPTIME[CMCwOPTIME$Complication=="FALSE",]$OPTIME)
sd(CMCwOPTIME[CMCwOPTIME$Complication=="TRUE",]$OPTIME)
sd(CMCwoComplication$OPTIME)
sd(CMCwithComplication$OPTIME)
range(CMCwoComplication$OPTIME)
range(CMCwithComplication$OPTIME)
```


HTOODAY check
```{r}
CMCwHTOODAY<-filter(CMCaddComplication, HTOODAY>=0)
CMCwHTOODAY$HTOODAY<-as.numeric(CMCwHTOODAY$HTOODAY)
dim(CMCwHTOODAY)
HTOODAYtt<-t.test(HTOODAY~Complication, data = CMCwHTOODAY)
HTOODAYtt$p.value
sd(CMCwHTOODAY[CMCwHTOODAY$Complication=="FALSE",]$HTOODAY)
sd(CMCwHTOODAY[CMCwHTOODAY$Complication=="TRUE",]$HTOODAY)

CMCwSTOODAY<-filter(CMCaddComplication, STOODAY>=0)
CMCwSTOODAY$STOODAY<-as.numeric(CMCwSTOODAY$STOODAY)
dim(CMCwSTOODAY)
STOODAYtt<-t.test(STOODAY~Complication, data = CMCwSTOODAY)
STOODAYtt$p.value
sd(CMCwoComplication$STOODAY)
sd(CMCwithComplication$STOODAY)
```


ANETIME
```{r}
CMCwANETIME<-filter(CMCaddComplication, ANETIME>=0)
CMCwANETIME$ANETIME<-as.numeric(CMCwANETIME$ANETIME)
dim(CMCwANETIME)
ANETIMEtt<-t.test(ANETIME~Complication, data = CMCwANETIME)
ANETIMEtt$p.value
ANETIMEtt
sd(CMCwANETIME[CMCwANETIME$Complication=="FALSE",]$ANETIME)
sd(CMCwANETIME[CMCwANETIME$Complication=="TRUE",]$ANETIME)
```




-----------------
http://www.stats.uwo.ca/faculty/murdoch/ism2013/4tables.pdf
https://www.inf.ethz.ch/personal/markusp/teaching/guides/guide-tables.pdf
https://www.r-bloggers.com/making-publication-ready-tables-with-xtable/
SAS: http://www.lexjansen.com/nesug/nesug03/ad/ad001.pdf
LaTeX: http://tex.stackexchange.com/questions/112343/beautiful-table-samples
LaTeX: http://cpansearch.perl.org/src/LIMAONE/LaTeX-Table-v1.0.6/examples/examples.pdf
LaTeX: http://ctan.org/pkg/booktabs
-----------------

6. Multivariate logistic regression

```{r,echo=FALSE}
CMCaddComplication$AGE<-as.numeric(CMCaddComplication$AGE)
CMCaddComplication$Complication<-factor(CMCaddComplication$Complication, levels = c("FALSE", "TRUE"))
CMCaddComplication$DIABETES<-factor(CMCaddComplication$DIABETES, levels = c("NO", "INSULIN", "ORAL", "NON-INSULIN", "NULL"))
fit1<-glm(Complication~DIABETES+ASACLAS+DIALYSIS+sixtyfive, data = CMCaddComplication, family = binomial)
summary(fit1)
exp(coef(fit1))
exp(cbind(OR = coef(fit1), confint(fit1)))
#######Interpreting Odds ratio: https://select-statistics.co.uk/blog/analysing-categorical-data-using-logistic-regression-models/


#http://stackoverflow.com/questions/18449013/r-logistic-regression-area-under-curve
prob=predict(fit1,type=c("response"))
CMCwithPred<-CMCaddComplication
CMCwithPred$prob<-prob
library(pROC)
g <- roc(Complication ~ prob, data = CMCwithPred)
plot(g) 
ci(g)
```
