---
title: "CMC"
author: "Bo Wang"
date: "November 20, 2016"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, include=FALSE}
library(sas7bdat)
library(dplyr)
library(plyr)
library(ggplot2)
library("fmsb")
```

0. NSQIP: download automatically 

```{r, echo=FALSE, include=FALSE}
load("nsqip05_06.RDA")
load("nsqip07.RDA")
load("nsqip08.RDA")
load("nsqip09.RDA") 
load("nsqip10.RDA")
load("nsqip11.RDA")
load("nsqip12.RDA")
load("nsqip13.RDA")
load("nsqip14.RDA")
load("nsqip15.RDA")
```


1. Data: CMC-Interposition arthroplasty, intercarpal or carpometacarpal joints (25447) and Arthroplasty with prosthetic replacement, trapezium (25445). 


```{r, echo=FALSE, include=FALSE}
### Pull out all Interposition arthroplasty, intercarpal or carpometacarpal joints (25447) from one year's dataset. It's easier to keep each year's data separate since they include different number of variables under different names. Combining all years' datasets after subsetting will save computing power.  
intArthFun<-function(data){
  temp<-data %>%
  filter(CPT==25447)
  intArthYr<-temp %>%
    arrange(CPT)
}

intArth05_06<-intArthFun(nsqip05_06)
intArth07<-intArthFun(nsqip07)
intArth08<-intArthFun(nsqip08)
intArth09<-intArthFun(nsqip09)
intArth10<-intArthFun(nsqip10)
intArth11<-intArthFun(nsqip11)
intArth12<-intArthFun(nsqip12)
intArth13<-intArthFun(nsqip13)
intArth14<-intArthFun(nsqip14)
intArth15<-intArthFun(nsqip15)
```


```{r, echo=FALSE, include=FALSE}
### Pull out all Arthroplasty with prosthetic replacement, trapezium (25445) from one dataset
ArthProFun<-function(data){
  temp<-data %>%
  filter(CPT==25445)
  ArthProYr<-temp %>%
    arrange(CPT)
}

ArthPro05_06<-ArthProFun(nsqip05_06)
ArthPro07<-ArthProFun(nsqip07)
ArthPro08<-ArthProFun(nsqip08)
ArthPro09<-ArthProFun(nsqip09)
ArthPro10<-ArthProFun(nsqip10)
ArthPro11<-ArthProFun(nsqip11)
ArthPro12<-ArthProFun(nsqip12)
ArthPro13<-ArthProFun(nsqip13)
ArthPro14<-ArthProFun(nsqip14)
ArthPro15<-ArthProFun(nsqip15)
```


```{r, echo=FALSE, include=FALSE}
### CMCyr is a combiniation of both intercarpal or carpometacarpal joints (25447) and arthroplasty with prosthetic replacement, trapezium (25445)
CMC05_06<-rbind(intArth05_06, ArthPro05_06)
CMC07<-rbind(intArth07, ArthPro07)
CMC08<-rbind(intArth08, ArthPro08)
CMC09<-rbind(intArth09, ArthPro09)
CMC10<-rbind(intArth10, ArthPro10)
CMC11<-rbind(intArth11, ArthPro11)
CMC12<-rbind(intArth12, ArthPro12)
CMC13<-rbind(intArth13, ArthPro13)
CMC14<-rbind(intArth14, ArthPro14)
CMC15<-rbind(intArth15, ArthPro15)

### Change all column names to UPPER in preparation of binding
names(CMC05_06)<-toupper(names(CMC05_06))
names(CMC07)<-toupper(names(CMC07))
names(CMC08)<-toupper(names(CMC08))
names(CMC09)<-toupper(names(CMC09))
names(CMC10)<-toupper(names(CMC10))
names(CMC11)<-toupper(names(CMC11))
names(CMC12)<-toupper(names(CMC12))
names(CMC13)<-toupper(names(CMC13))
names(CMC14)<-toupper(names(CMC14))
names(CMC15)<-toupper(names(CMC15))
```


```{r, echo=FALSE, include=FALSE}
### Binding
### http://stackoverflow.com/questions/3402371/combine-two-data-frames-by-rows-rbind-when-they-have-different-sets-of-columns

### RACE for 2005-07, RACE_NEW 2008-15
names(CMC08)[3]<-"RACE"
names(CMC09)[3]<-"RACE"
names(CMC10)[3]<-"RACE"
names(CMC11)[3]<-"RACE"
names(CMC12)[3]<-"RACE"
names(CMC13)[3]<-"RACE"
names(CMC14)[3]<-"RACE"
names(CMC15)[4]<-"RACE"

CMC<-rbind.fill(CMC05_06, CMC07, CMC08, CMC09, CMC10, CMC11, CMC12, CMC13, CMC14, CMC15)

### RACE 2005-08 change to RACE_NEW values
CMC$RACE<-gsub("White,(.*)?", "White", CMC$RACE)
CMC$RACE<-gsub("Hispanic, White", "White", CMC$RACE)
CMC$RACE<-gsub("Hispanic, Black", "Black or African American", CMC$RACE)
CMC$RACE<-gsub("Unknown(.*)?", "Unknown", CMC$RACE)
CMC$RACE<-gsub("NULL", "Unknown", CMC$RACE)

### AGE to numeric
which(CMC$AGE=="90+")
### https://stat.ethz.ch/pipermail/r-help/2008-December/182748.html
levels(CMC$AGE) <- c(levels(CMC$AGE), "90")
CMC$AGE[c(473,1642,2214)]<-c("90", "90", "90")
CMC$AGE<-as.numeric(as.character(CMC$AGE))
```


2. Basic summary of CMC data

Number of patients and variables overall
```{r}
dim(CMC)
```

Number of patients and variables for each year
```{r, echo=FALSE}
dim05_06<-dim(CMC05_06)
dim07<-dim(CMC07)
dim08<-dim(CMC08)
dim09<-dim(CMC09)
dim10<-dim(CMC10)
dim11<-dim(CMC11)
dim12<-dim(CMC12)
dim13<-dim(CMC13)
dim14<-dim(CMC14)
dim15<-dim(CMC15)
dims<-c(dim05_06, dim07, dim08, dim09, dim10, dim11, dim12, dim13, dim14, dim15)
dims<-as.data.frame(matrix(dims, nrow = 2))
rownames(dims)<-c("patients", "variables")
colnames(dims)<-c("2005-06", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")
dims
```

Summary tables
```{r, echo=FALSE}
tables<-apply(CMC, 2, table)
tables["RACE"]
tables["SEX"]
tables["CPT"]
```

See specific table(s)
```{r}
which(names(CMC)=="RACE")
tables[41:42]
```

Summaries as data frame
```{r, echo=FALSE}
### see tables in df form
tbToDf<-function(tb){
  df<-as.data.frame(unlist(tb))
  names(df)<-c("","Count")
  df
}
dfs<-lapply(tables, tbToDf)
```

```{r}
names(dfs)<-names(CMC)
dfs["RACE"]
dfs["SEX"]
```


3. Complications
```{r}
complications<-c("SUPINFEC", "SSSIPATOS", "WNDINFD", 
        "DSSIPATOS", "ORGSPCSSI", "OSSIPATOS",
        "DEHIS", "OUPNEUMO", "PNAPATOS",    
        "REINTUB", "PULEMBOL","FAILWEAN", 
        "VENTPATOS", "RENAINSF", "OPRENAFL",
        "URNINFEC", "UTIPATOS",  "CNSCVA",
        "CDARREST", "CDMI", "OTHBLEED",  
        "OTHDVT", "OTHSYSEP",  "SEPSISPATOS", 
        "OTHSESHOCK", "SEPSHOCKPATOS", "RETURNOR",
        "OTHCDIFF", "RETORRELATED", "RETOR2RELATED", 
        "READMRELATED1", "READMRELATED2", "READMRELATED3", 
        "READMRELATED4", "READMRELATED5", "REOPERATION",
        "READMISSION", "UNPLANREADMISSION", "OTHGRAFL", 
        "COMA", "NEURODEF")
```

Number of unique patients with complications (1.6%)
```{r, echo=FALSE}
### function to assign Y/N complication column 
complicationOrNot<-function(df){
  Complication<-vector()
for(i in 1:nrow(df)) {if((all(df[,complications][i,]=="No" 
                          | df[,complications][i,]=="No Complication"
                          | df[,complications][i,]=="NULL"
                          | is.na(df[,complications][i,])))==TRUE){
                 Complication[i]<-"FALSE"
                 } else {
                 Complication[i]<-"TRUE"}}
  df<-cbind(df, Complication)
}
CMCaddComplication<-complicationOrNot(CMC)
CMCaddComplication[CMC$DOPERTOD==29,]$Complication<-TRUE
table(CMCaddComplication$Complication)
```


0.3% Return OR 
```{r, echo=FALSE}
table(CMCaddComplication$RETURNOR)
```

Distribution of complications (SAME AS DRfx PAPER)
```{r, echo=FALSE}
CMCwithComplication<-CMCaddComplication[CMCaddComplication$Complication==TRUE,]
CMCwoComplication<-CMCaddComplication[CMCaddComplication$Complication==FALSE,]
complicationTables<-list(
table(CMCwithComplication$SUPINFEC),
table(CMCwithComplication$WNDINFD),
table(CMCwithComplication$ORGSPCSSI),
table(CMCwithComplication$DEHIS),
table(CMCwithComplication$OUPNEUMO),
table(CMCwithComplication$PULEMBOL),
table(CMCwithComplication$REINTUB),
table(CMCwithComplication$FAILWEAN),
table(CMCwithComplication$CDARREST),
table(CMCwithComplication$CDMI),
table(CMCwithComplication$RENAINSF),
table(CMCwithComplication$OPRENAFL),
table(CMCwithComplication$CNSCVA),
table(CMCwithComplication$OTHGRAFL),
table(CMCwithComplication$COMA),
table(CMCwithComplication$NEURODEF),
table(CMCwithComplication$OTHDVT),
table(CMCwithComplication$OTHSYSEP),
table(CMCwithComplication$OTHSESHOCK),
table(CMCwithComplication$RETURNOR),
table(CMCwithComplication$OTHBLEED),
table(CMCwithComplication$URNINFEC)
)



compl1<-c("SUPINFEC", "WNDINFD", 
        "ORGSPCSSI",
        "DEHIS", "OUPNEUMO","PULEMBOL", 
        "REINTUB", "FAILWEAN", 
        "CDARREST","CDMI", "RENAINSF", "OPRENAFL",
        "CNSCVA", "OTHGRAFL",
        "COMA","NEURODEF", "OTHDVT",
        "OTHSYSEP", "OTHSESHOCK",
        "RETURNOR", "OTHBLEED",
        "URNINFEC")
complicationdfs<-lapply(complicationTables, tbToDf)
names(complicationdfs)<-compl1
```

```{r, echo=FALSE, eval=FALSE}
complicationdfs[1:20]
```



Distribution of complications (Added Readmission, OTHCDIFF, etc.)
```{r, echo=FALSE}
CMCwithComplication<-CMCaddComplication[CMCaddComplication$Complication==TRUE,]
complicationTables2<-list(
table(CMCwithComplication$SUPINFEC),
table(CMCwithComplication$WNDINFD),
table(CMCwithComplication$ORGSPCSSI),
table(CMCwithComplication$DEHIS),
table(CMCwithComplication$OUPNEUMO),
table(CMCwithComplication$PULEMBOL),
table(CMCwithComplication$REINTUB),
table(CMCwithComplication$FAILWEAN),
table(CMCwithComplication$CDARREST),
table(CMCwithComplication$CDMI),
table(CMCwithComplication$RENAINSF),
table(CMCwithComplication$OPRENAFL),
table(CMCwithComplication$CNSCVA),
table(CMCwithComplication$OTHGRAFL),
table(CMCwithComplication$COMA),
table(CMCwithComplication$NEURODEF),
table(CMCwithComplication$OTHDVT),
table(CMCwithComplication$OTHSYSEP),
table(CMCwithComplication$OTHSESHOCK),
table(CMCwithComplication$RETURNOR),
table(CMCwithComplication$OTHBLEED),
table(CMCwithComplication$URNINFEC),

table(CMCwithComplication$SSSIPATOS),
table(CMCwithComplication$DSSIPATOS),
table(CMCwithComplication$OSSIPATOS),
table(CMCwithComplication$PNAPATOS),
table(CMCwithComplication$VENTPATOS),
table(CMCwithComplication$UTIPATOS),
table(CMCwithComplication$SEPSISPATOS),
table(CMCwithComplication$SEPSHOCKPATOS),

table(CMCwithComplication$OTHCDIFF),
table(CMCwithComplication$RETORRELATED),
table(CMCwithComplication$RETOR2RELATED),
table(CMCwithComplication$READMRELATED1),
table(CMCwithComplication$READMRELATED2),
table(CMCwithComplication$READMRELATED3),
table(CMCwithComplication$READMRELATED4),
table(CMCwithComplication$READMRELATED5),
table(CMCwithComplication$REOPERATION),
table(CMCwithComplication$READMISSION),
table(CMCwithComplication$UNPLANREADMISSION)
)
complicationdfs2<-lapply(complicationTables2, tbToDf)
compl2<-c("SUPINFEC", "WNDINFD", 
        "ORGSPCSSI",
        "DEHIS", "OUPNEUMO","PULEMBOL", 
        "REINTUB", "FAILWEAN", 
        "CDARREST","CDMI", "RENAINSF", "OPRENAFL",
        "CNSCVA", "OTHGRAFL",
        "COMA","NEURODEF", "OTHDVT",
        "OTHSYSEP", "OTHSESHOCK",
        "RETURNOR", "OTHBLEED",
        "URNINFEC", "SSSIPATOS",   
        "DSSIPATOS", "OSSIPATOS",
        "PNAPATOS", "VENTPATOS",   
        "UTIPATOS", "SEPSISPATOS", 
        "SEPSHOCKPATOS", "OTHCDIFF",
        "RETORRELATED", "RETOR2RELATED", 
                 "READMRELATED1", "READMRELATED2", "READMRELATED3", 
                 "READMRELATED4", "READMRELATED5", "REOPERATION",
                 "READMISSION", "UNPLANREADMISSION"
        )
names(complicationdfs2)<-compl2

####################################################################
#crtTbLst<-function(df, vec){                                      #
#complicationdfs2<-list()                                          #
#for(i in 1:length(vec)){                                          #
#  complicationdfs2[i]<-table(df[,vec[i]])                         #
#}                                                                 #
#complicationdfs2                                                  #
#}                                                                 #
#                                                                  #
#complicationTables3<-crtTbLst(CMCwithComplication, compl2)        #
####################################################################

```

Total of 72 complications
```{r, echo=FALSE}
complicationdfs2[1:41]
```


4. 2*2 Tables (categorical variable v. Complication (Y/N))

```{r}
##function to get rid of all 0 rows in a table
ridZero<-function(tb){
  num<-1
  while(num<=dim(tb)[1]){
    if(all(tb[num,]==0)==TRUE){
      tb<-tb[-num,]
      num<-num
    }else{
      tb<-tb
      num<-num+1
    }
  }
  tb
}
```

SEX
```{r, echo=FALSE}
SEXtb<-table(CMCaddComplication[, c("SEX", "Complication")])
SEXtb
SEXtb<-SEXtb[-3,]
SEXtb
SEXchisq<-chisq.test(SEXtb, simulate.p.value = TRUE)
SEXchisq$p.value
```

RACE
```{r, echo=FALSE}
RACEtb<-table(CMCaddComplication[, c("RACE", "Complication")])
RACEtb
RACEtb<-RACEtb[-5,]
RACEtb
RACEchisq<-chisq.test(RACEtb, simulate.p.value = TRUE)
RACEchisq$p.value
RACEanova<-anova(RACEtb)
```

INOUT
```{r, echo=FALSE}
INOUTtb<-table(CMCaddComplication[, c("INOUT", "Complication")])
INOUTtb<-INOUTtb[-3,]
INOUTtb
INOUTchisq<-chisq.test(INOUTtb, simulate.p.value = TRUE)
INOUTchisq$p.value
```

TRANST
```{r, echo=FALSE}
TRANSTtb<-table(CMCaddComplication[, c("TRANST", "Complication")])
TRANSTtb<-ridZero(TRANSTtb)
TRANSTtb
TRANSTchisq<-chisq.test(TRANSTtb, simulate.p.value = TRUE)
TRANSTchisq$p.value

TRANSTtb2<-TRANSTtb[c(1,4),]
TRANSTtb2
TRANSTchisq2<-chisq.test(TRANSTtb2, simulate.p.value = TRUE)
TRANSTchisq2$p.value
```

ATTEND
```{r, echo=FALSE}
ATTENDtb<-table(CMCaddComplication[, c("ATTEND", "Complication")])
ATTENDtb<-ridZero(ATTENDtb)
ATTENDtb
ATTENDtb2<-ATTENDtb[c(-3,-5),]
ATTENDtb2
ATTENDchisq2<-chisq.test(ATTENDtb2, simulate.p.value = TRUE)
ATTENDchisq2$p.value
```

ANESTHES
```{r, echo=FALSE}
ANESTHEStb<-table(CMCaddComplication[, c("ANESTHES", "Complication")])
ANESTHEStb<-ridZero(ANESTHEStb)
ANESTHEStb
ANESTHESchisq<-chisq.test(ANESTHEStb, simulate.p.value = TRUE)
ANESTHESchisq$p.value
ANESTHEStb[,2]/(ANESTHEStb[,1]+ANESTHEStb[,2])
ANESTHEStb2<-ANESTHEStb[c(2,7,11),]
ANESTHEStb2
ANESTHESchisq2<-chisq.test(ANESTHEStb2, simulate.p.value = TRUE)
ANESTHESchisq2$p.value
ANESTHEStb2[,2]/(ANESTHEStb2[,1]+ANESTHEStb2[,2])
```

SURGSPEC
```{r, echo=FALSE}
SURGSPECtb<-table(CMCaddComplication[, c("SURGSPEC", "Complication")])
SURGSPECtb<-ridZero(SURGSPECtb)
SURGSPECtb
SURGSPECchisq<-chisq.test(SURGSPECtb, simulate.p.value = TRUE)
SURGSPECchisq$p.value
SURGSPECtb2<-SURGSPECtb[c(1,3,5),]
SURGSPECtb2
SURGSPECchisq2<-chisq.test(SURGSPECtb2, simulate.p.value = TRUE)
SURGSPECchisq2$p.value
```

DIABETES
```{r, echo=FALSE}
DIABETEStb<-table(CMCaddComplication[, c("DIABETES", "Complication")])
DIABETEStb<-ridZero(DIABETEStb)
DIABETEStb
DIABETESchisq<-chisq.test(DIABETEStb, simulate.p.value = TRUE)
DIABETESchisq$p.value
DIABETEStb[,2]/(DIABETEStb[,1]+DIABETEStb[,2])
```

SMOKE
```{r, echo=FALSE}
SMOKEtb<-table(CMCaddComplication[, c("SMOKE", "Complication")])
SMOKEtb<-ridZero(SMOKEtb)
SMOKEtb
SMOKEchisq<-chisq.test(SMOKEtb, simulate.p.value = TRUE)
SMOKEchisq$p.value
```

ETOH
```{r, echo=FALSE}
ETOHtb<-table(CMCaddComplication[, c("ETOH", "Complication")])
ETOHtb<-ridZero(ETOHtb)
ETOHtb<-ETOHtb[-3,]
ETOHtb
ETOHchisq<-chisq.test(ETOHtb, simulate.p.value = TRUE)
ETOHchisq$p.value
```

CPT
```{r, echo=FALSE}
CPTtb<-table(CMCaddComplication[, c("CPT", "Complication")])
CPTtb<-ridZero(CPTtb)
CPTtb
CPTchisq<-chisq.test(CPTtb, simulate.p.value = TRUE)
CPTchisq$p.value
```

HYPERMED
```{r, echo=FALSE}
HYPERMEDtb<-table(CMCaddComplication[, c("HYPERMED", "Complication")])
HYPERMEDtb<-ridZero(HYPERMEDtb)
HYPERMEDtb
HYPERMEDchisq<-chisq.test(HYPERMEDtb, simulate.p.value = TRUE)
HYPERMEDchisq$p.value
```

WNDCLAS
```{r, echo=FALSE}
WNDCLAStb<-table(CMCaddComplication[, c("WNDCLAS", "Complication")])
WNDCLAStb<-ridZero(WNDCLAStb)
WNDCLAStb
WNDCLASchisq<-chisq.test(WNDCLAStb, simulate.p.value = TRUE)
WNDCLASchisq$p.value
```

ASACLAS
```{r, echo=FALSE}
ASACLAStb<-table(CMCaddComplication[, c("ASACLAS", "Complication")])
ASACLAStb<-ridZero(ASACLAStb)
ASACLAStb<-ASACLAStb[-5,]
ASACLAStb
ASACLASchisq<-chisq.test(ASACLAStb, simulate.p.value = TRUE)
ASACLASchisq$p.value
ASACLAStb[,2]/(ASACLAStb[,1]+ASACLAStb[,2])
```

BLEEDDIS
```{r, echo=FALSE}
BLEEDDIStb<-table(CMCaddComplication[, c("BLEEDDIS", "Complication")])
BLEEDDIStb<-ridZero(BLEEDDIStb)
BLEEDDIStb
BLEEDDISchisq<-chisq.test(BLEEDDIStb, simulate.p.value = TRUE)
BLEEDDISchisq$p.value
BLEEDDIStb[,2]/(BLEEDDIStb[,1]+BLEEDDIStb[,2])
```

HXCHF
```{r, echo=FALSE}
HXCHFtb<-table(CMCaddComplication[, c("HXCHF", "Complication")])
HXCHFtb<-ridZero(HXCHFtb)
HXCHFtb
HXCHFchisq<-chisq.test(HXCHFtb, simulate.p.value = TRUE)
HXCHFchisq$p.value
```

STEROID
```{r, echo=FALSE}
STEROIDtb<-table(CMCaddComplication[, c("STEROID", "Complication")])
STEROIDtb<-ridZero(STEROIDtb)
STEROIDtb
STEROIDchisq<-chisq.test(STEROIDtb, simulate.p.value = TRUE)
STEROIDchisq$p.value
```

CHEMO
```{r, echo=FALSE}
CHEMOtb<-table(CMCaddComplication[, c("CHEMO", "Complication")])
CHEMOtb<-ridZero(CHEMOtb)
CHEMOtb<-CHEMOtb[-3,]
CHEMOtb
CHEMOchisq<-chisq.test(CHEMOtb, simulate.p.value = TRUE)
CHEMOchisq$p.value
```

RADIO
```{r, echo=FALSE}
RADIOtb<-table(CMCaddComplication[, c("RADIO", "Complication")])
RADIOtb<-ridZero(RADIOtb)
RADIOtb<-RADIOtb[-3,]
RADIOtb
RADIOchisq<-chisq.test(RADIOtb, simulate.p.value = TRUE)
RADIOchisq$p.value
```

PRVPCI (all but year 2015)
```{r, echo=FALSE}
PRVPCItb<-table(CMCaddComplication[, c("PRVPCI", "Complication")])
PRVPCItb<-ridZero(PRVPCItb)
PRVPCItb<-PRVPCItb[-3,]
PRVPCItb
PRVPCIchisq<-chisq.test(PRVPCItb, simulate.p.value = TRUE)
PRVPCIchisq$p.value
```

TRANSFUS
```{r, echo=FALSE}
TRANSFUStb<-table(CMCaddComplication[, c("TRANSFUS", "Complication")])
TRANSFUStb
```



Predictors I didn't find in dataset (PUF):

Intra-articular fracture

Associated ORIF ulna styloid

Current alcohol use

Preoperative function - FNSTATUS1/FNSTATUS2
All year but 2015 has both variables, 2015 only has FNSTATUS2
```{r}
table(CMCaddComplication$FNSTATUS1)
```

Chronic pulmonary obstructive disease

Peripheral vascular disease

Cardiac disease

Cerebrovascular accident or transient ischemic attack



5. Logistic regression (Continuous variable v. Complication (Y/N)))

WORKRVU
```{r, echo=FALSE}
WORKRVUtt<-t.test(WORKRVU~Complication, data = CMCaddComplication)
WORKRVUtt$p.value
```

AGE
```{r}
AGEtt<-t.test(AGE~Complication, data = CMCaddComplication)
AGEtt
sd(CMCwoComplication$OPTIME)
range(CMCwoComplication$OPTIME)
sd(CMCwithComplication$OPTIME)
range(CMCwithComplication$OPTIME)
```

```{r, include=FALSE}
overForty<-function(df){
  Forty<-vector()
for(i in 1:nrow(df)) {
  if(df$AGE[i]>40){
  Forty[i]<-TRUE
  } else if(df$AGE[i]<=40){
  Forty[i]<-FALSE
  }
  }
  df<-cbind(df, Forty)
  df
}
#####################################RISK UNDER/ABOVE 40##########################################
```

BMI
```{r, echo=FALSE}
CMCwHeight<-filter(CMCaddComplication, HEIGHT>=0)
CMCwHeightWeight<-filter(CMCwHeight, WEIGHT>=0)
dim(CMCwHeightWeight)
CMCwHeightWeight$BMI<-CMCwHeightWeight$WEIGHT*0.45/(CMCwHeightWeight$HEIGHT*0.025)^2

BMItt<-t.test(BMI~Complication, data = CMCwHeightWeight)
BMItt$p.value
```

PRPT
```{r}
CMCwPRPT<-CMCaddComplication[CMCaddComplication$PRPT!=-99,]
dim(CMCwPRPT)
table(CMCwPRPT$Complication)
```

PRINR
```{r}
CMCwPRINR<-CMCaddComplication[CMCaddComplication$PRINR!=-99,]
dim(CMCwPRINR)
table(CMCwPRINR$Complication)
PRINRtt<-t.test(PRINR~Complication, data = CMCwPRINR)
PRINRtt$p.value
```

PRWBC
```{r}
CMCwPRWBC<-filter(CMCaddComplication, PRWBC>=0)
dim(CMCwPRWBC)
PRWBCtt<-t.test(PRWBC~Complication, data = CMCwPRWBC)
PRWBCtt$p.value
```

PRCREAT
```{r}
CMCwPRCREAT<-filter(CMCaddComplication, PRCREAT>=0)
dim(CMCwPRCREAT)
PRCREATtt<-t.test(PRCREAT~Complication, data = CMCwPRCREAT)
PRCREATtt$p.value
```

PRHCT
```{r}
CMCwPRHCT<-filter(CMCaddComplication, PRHCT>=0)
dim(CMCwPRHCT)
PRHCTtt<-t.test(PRHCT~Complication, data = CMCwPRHCT)
PRHCTtt
```

OPTIME
```{r}
CMCwOPTIME<-filter(CMCaddComplication, OPTIME>=0)
dim(CMCwOPTIME)
OPTIMEtt<-t.test(OPTIME~Complication, data = CMCwOPTIME)
OPTIMEtt$p.value
sd(CMCwoComplication$AGE)
range(CMCwoComplication$AGE)
sd(CMCwithComplication$AGE)
range(CMCwithComplication$AGE)
```


-----------------
http://www.stats.uwo.ca/faculty/murdoch/ism2013/4tables.pdf
https://www.inf.ethz.ch/personal/markusp/teaching/guides/guide-tables.pdf
https://www.r-bloggers.com/making-publication-ready-tables-with-xtable/
SAS: http://www.lexjansen.com/nesug/nesug03/ad/ad001.pdf
LaTeX: http://tex.stackexchange.com/questions/112343/beautiful-table-samples
LaTeX: http://cpansearch.perl.org/src/LIMAONE/LaTeX-Table-v1.0.6/examples/examples.pdf
LaTeX: http://ctan.org/pkg/booktabs
-----------------


```{r, eval=FALSE, echo=FALSE}
complications11<-c("ADMYR", "CASEID", "SUPINFEC", "SSSIPATOS", "WNDINFD", 
        "DSSIPATOS", "ORGSPCSSI", "OSSIPATOS",
        "DEHIS", "OUPNEUMO", "PNAPATOS",    
        "REINTUB", "PULEMBOL","FAILWEAN", 
        "VENTPATOS", "RENAINSF", "OPRENAFL",
        "URNINFEC", 
        ## MIGHT DELETE PATOS
        "UTIPATOS",  
        "CNSCVA",
        "CDARREST", "CDMI", "OTHBLEED",  
        "OTHDVT", "OTHSYSEP", 
        "OTHSESHOCK", "RETURNOR", "REOPERATION", 
        "READMISSION", "UNPLANREADMISSION")

complications12<-c("ADMYR","CASEID", "SUPINFEC", "SSSIPATOS", "WNDINFD", 
        "DSSIPATOS", "ORGSPCSSI", "OSSIPATOS",
        "DEHIS", "OUPNEUMO", "PNAPATOS",    
        "REINTUB", "PULEMBOL","FAILWEAN", 
        "VENTPATOS", "RENAINSF", "OPRENAFL",
        "URNINFEC", "UTIPATOS",  "CNSCVA",
        "CDARREST", "CDMI", "OTHBLEED",  
        "OTHDVT", "OTHSYSEP",  "SEPSISPATOS", 
        "OTHSESHOCK", "SEPSHOCKPATOS", "RETURNOR", 
        "RETORRELATED", "RETOR2RELATED","READMRELATED1", 
        "READMRELATED2", "READMRELATED3", "READMRELATED4", 
        "READMRELATED5")

complications13<-c("AdmYR","CaseID", "SUPINFEC", "SSSIPATOS", "WNDINFD", 
        "DSSIPATOS", "ORGSPCSSI", "OSSIPATOS",
        "DEHIS", "OUPNEUMO", "PNAPATOS",    
        "REINTUB", "PULEMBOL","FAILWEAN", 
        "VENTPATOS", "RENAINSF", "OPRENAFL",
        "URNINFEC", "UTIPATOS",  "CNSCVA",
        "CDARREST", "CDMI", "OTHBLEED",  
        "OTHDVT", "OTHSYSEP",  "SEPSISPATOS", 
        "OTHSESHOCK", "SEPSHOCKPATOS", "RETURNOR",
        "RETORRELATED", "RETOR2RELATED","READMRELATED1", 
        "READMRELATED2", "READMRELATED3", "READMRELATED4", "READMRELATED5")

complications14<-c("CaseID", "SUPINFEC", "SSSIPATOS", "WNDINFD", 
        "DSSIPATOS", "ORGSPCSSI", "OSSIPATOS",
        "DEHIS", "OUPNEUMO", "PNAPATOS",    
        "REINTUB", "PULEMBOL","FAILWEAN", 
        "VENTPATOS", "RENAINSF", "OPRENAFL",
        "URNINFEC", "UTIPATOS",  "CNSCVA",
        "CDARREST", "CDMI", "OTHBLEED",  
        "OTHDVT", "OTHSYSEP",  "SEPSISPATOS", 
        "OTHSESHOCK", "SEPSHOCKPATOS", "RETURNOR",
        "RETORRELATED", "RETOR2RELATED","READMRELATED1", 
        "READMRELATED2", "READMRELATED3", "READMRELATED4", "READMRELATED5")

### http://stackoverflow.com/questions/11302985/match-with-negation
### Negation of %in% -- no in
"%ni%"<-Negate("%in%")
complications11[complications11 %ni% complications]
complications12[complications12 %ni% complications]
complications13[complications13 %ni% complications]
complications14[complications14 %ni% complications]
```

Here are all variables not included as "complications" 
```{r, echo=FALSE}
"%ni%"<-Negate("%in%")
names(CMC)[names(CMC) %ni% complications]
```










